\section{Interpretation and Synthesis}

\subsection{Defending the System Property Claim}
The central thesis of this work is that \textbf{reliability is a system property, not a component property}. Our empirical results support this through three key observations:
\begin{enumerate}
    \item \textbf{Individual Volatility}: Components like resENC and resTR are not inherently "safe." They function correctly within narrow regimes but propagate or amplify errors once input assumptions are violated.
    \item \textbf{Emergent Safety}: The system's ability to "refuse" to generate is not programmed into the decoder, nor is it a feature of the encoder. It is an emergent outcome of the interaction between the statistical sensors (observability) and the control surface (governance).
    \item \textbf{Structural Generalization}: By utilizing structural calibration (Z-mapping), we show that reliability logic remains valid even when the data geometry changes radically (e.g., from vision to biology).
\end{enumerate}

\subsection{Transparency vs. Robustness}
Conventional research often pursues "robust models" that attempt to generalize to all possible corner cases. resED takes the opposite approach: it assumes components \textit{will} fail and prioritizes **transparency**. By converting opaque latent states into explicit risk coordinates, we enable a system that is predictable and auditable, even when its constituent models are black boxes.

\section{Limitations and Non-Claims}
To maintain scientific rigor, we explicitly define the boundaries of the resED system:
\begin{itemize}
    \item \textbf{Accuracy vs. Reliability}: resED does not fix incorrect encodings. It only identifies and blocks them. It is a "fail-safe" system, not an "error-correcting" system.
    \item \textbf{Adversarial Bounds}: While we have tested stochastic noise and structured shifts, we make no formal claims regarding robustness against optimized adversarial attacks designed to hide in the manifold's high-probability regions.
    \item \textbf{Semantic Blindness}: The system is semantically agnostic. It judges representations by their statistical "typicality," not their content.
\end{itemize}

\section{Conclusion}
We have presented resED, a governed encoder-decoder framework that enforces reliability through representation-level control. By implementing a modular architecture consisting of deterministic encoding, statistical sensing, and gated refinement, we have demonstrated a system capable of detecting and mitigating latent failure across diverse domains. Our findings suggest that for safety-critical applications, the engineering of the governance layer is as important as the optimization of the generative backbone.