\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{letterpaper, margin=1in}

\begin{document}
\thispagestyle{empty}

\noindent \today

\vspace{0.5cm}

\noindent \textbf{To the Editor-in-Chief,} \\
\textbf{IEEE Transactions on Neural Networks and Learning Systems (TNNLS)}

\vspace{0.5cm}

\noindent \textbf{Subject: Submission of Manuscript "Reliability is a System Property"}

\vspace{0.5cm}

\noindent Dear Editor,

I am pleased to submit the enclosed manuscript entitled "Reliability is a System Property" for consideration for publication in IEEE Transactions on Neural Networks and Learning Systems.

This paper proposes a paradigm shift in AI reliability: rather than attempting to harden individual deep learning components against all possible failures, we introduce a system-level governance architecture, \textbf{resED}, which wraps opaque generative models in a deterministic "Representation-Level Control Surface" (RLCS). Our core contribution is demonstrating that reliability can be guaranteed as an architectural property—specifically, the ability to fail safely and observably—even when using volatile, black-box components.

To ensure the highest standards of transparency and reproducibility, this work:
\begin{enumerate}
    \item Explicitly scopes its claims to exclude correctness or adversarial robustness guarantees, focusing solely on failure observability and containment.
    \item Highlights critical architectural limitations, specifically "Normalization Blindness" in LayerNorm-based Transformers.
    \item Provides a full reference implementation of the governance logic via the \texttt{resLIK} R package (available on CRAN) and makes the complete codebase available on GitHub.
\end{enumerate}

We believe this work addresses a critical gap in the safe deployment of generative AI by providing a practical, scalable mechanism for "circuit-breaking" model hallucinations. This work does not compete with OOD detectors; it formalizes when such detectors should be trusted.

None of the authors have a conflict of interest to disclose.

Thank you for your consideration.

\vspace{1cm}

\noindent Sincerely,

\vspace{0.5cm}

\noindent \textbf{MD. Arshad} \\
Department of Computer Science \\
Jamia Millia Islamia \\
arshad10867c@gmail.com

\end{document}
